---
title: "Week06b Lab: Making a Gorgeous Markdown Script"
author: "Hannah Merges"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty:
    theme: hpstr
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning=FALSE, 
                      message=FALSE, 
                      fig.path = "../Outputs/")
```
## Introduction 
Welcome! This lab will use R Markdown to: 

* Create a table and figure using a previous dataset 
* Use features such as **bold**, _italics_, and lists to demonstrate comfort in R Markdown language 
* Make a _beautiful script_ using new themes, document styles, texts, etc. 


### Load necessary libraries 
```{r}
library(tidyverse)
library(here)
library(kableExtra)
library(lubridate)
library(praise)
library(rmarkdown)
```

### Read in data and arrange/tidy it 

```{r}
DepthData <-read_csv(here("Week_06","Data","DepthData.csv"))

CondData<-read_csv(here("Week_06","Data","CondData.csv")) %>% 
  mutate(date=mdy_hms(date), ##change the columns to proper date, year format 
         date=round_date(date, "10 seconds")) #round seconds to match with depth data

###joining both datasets together to analyze in figures more easily 
joined_DepthCond <- inner_join(DepthData, CondData) %>% 
  mutate(date=round_date(date, "minute")) %>% #rounds for minute 
  group_by(date) %>% #group by the dates to then average everything 
  summarise(avg_temp=mean(Temperature), #averages for each var
            avg_salinity=mean(Salinity), 
            avg_depth=mean(Depth))

## will use this for the table 
full_DepthCond <- inner_join(DepthData, CondData)

totalaverages <- full_DepthCond %>% 
  group_by(Serial) %>% 
  summarise(totalavg_temp=mean(Temperature), #averages for each var
            totalavg_salinity=mean(Salinity), 
            totalavg_depth=mean(Depth))
totalaverages

```
### Make a figure using this data 

```{r conductivity_depth_figure, fig.align='center', fig.cap="The average temperature, salinity, and depth recordings from January 1st, 2015"}

##use stat_density_2d for this specific dataset because this uses the frequency of each point --> how often does this salinity occur at this temperature 
##can normally use geom_tile() for other heat maps 

joined_DepthCond %>%
  ggplot(aes(x=avg_salinity,
             y=avg_temp)) +
  stat_density_2d(aes(fill = ..level..), geom = "polygon", colour="white") + #area + contour 
  labs(x="Average Salinity (ppt)", 
       y="Average Temperature(C)", 
       title="Analysis of Temperature and Salinity Frequencies") + 
  theme_linedraw() +
  theme(axis.text.x=element_text(size=10), 
        axis.text.y=element_text(size=10), 
        axis.title.x=element_text(size=12),
        axis.title.y=element_text(size=12), 
        plot.title=element_text(hjust=0.5))
```

### Showing data in a table 
```{r}
totalaverages %>% 
  kbl() %>% #make it a kable table
  kable_classic() %>% #style for this kable table 
  row_spec(1, bold=T, color="white", background="coral") %>% #more style choices to change the color of my rows
  kable_styling(full_width=FALSE) #this makes the width fit to the size/amount of text 
```

### Summary 
This lab was designed to make a **beautiful** R Markdown script and practice new skills. I think I succeeded. I learned some new things like how to make a density plot/heat map. I also learned that the plots output right below in the script, not in the plots window like I am used to :)
Goals for next time: include pictures! 
**_Thank you for reading!_**













