---
title: "Week12a_Lab_ForLoops"
author: "Hannah Merges"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.path = "/..Outputs/")
```
 
## Assignment Goals 
For this lab, we will be working with for loops and the map function to calculate means and standard deviations using four different files from intertidal studies. 

## Load the necessary libraries 
```{r}
library(tidyverse)
library(here)
```

## Import the 4 datasets we will be using and create path for for loop  
```{r}
IntertidalPath<-here("Week_12", "Data", "homework")
IntertidalPath

tidal_files <- dir(path = IntertidalPath,pattern = ".csv")
tidal_files
```
## Make a blank df and pre-allocate space for R to run 
```{r}
# make an empty dataframe that has one row for each file and 5 columns (stdv and mean for light and temp and per tide pool)
intertidal_df <-data.frame(matrix(nrow = length(tidal_files), ncol = 5))
# give the dataframe column names
colnames(intertidal_df)<-c("tide_number","mean_temp", "stdv_temp", "mean_light", "stdv_light")
intertidal_df
```
## Test to make sure your for loop works then load all 4 
```{r}
raw_tidesdata<-read_csv(paste0(IntertidalPath,"/",tidal_files[1])) # test by reading in the first file and see if it works
head(raw_tidesdata)

## now load all 4 
all_tides_data<-read_csv(paste0(IntertidalPath,"/",tidal_files[1:4]))
head(all_tides_data)
```
## Calculate **means** and **standard deviation** for temperature and light using for loops 
```{r}
## mean temp first 
mean_temp<-mean(all_tides_data$Temp.C, na.rm = TRUE) # calculate a mean
mean_temp

for (i in 1:length(tidal_files)){ # loop over 1:3 the number of files 
  raw_data<-read_csv(paste0(IntertidalPath,"/",tidal_files[i]))
  #glimpse(raw_data)
  intertidal_df$tide_number[i]<-tidal_files[i] ##before the $ is whatever you named the df, after $ is first assigned column name with filenames  
  intertidal_df$mean_temp[i]<-mean(raw_data$Temp.C, na.rm =TRUE) # calculating mean
  intertidal_df$stdv_temp[i]<-sd(raw_data$Temp.C, na.rm =TRUE) # calculating stdv 
  intertidal_df$mean_light[i]<-mean(raw_data$Intensity.lux, na.rm =TRUE) # use raw.data, and the associated columns 
  intertidal_df$stdv_light[i]<-sd(raw_data$Intensity.lux, na.rm =TRUE)
}
intertidal_df

```
## Now do the same thing for mean and standard deviation for temperature and light but with _maps()_
```{r}
IntertidalPath_maps<-here("Week_12", "Data", "homework")
tides_files <- dir(path = IntertidalPath_maps,pattern = ".csv", full.names = TRUE) ## list all the files with .csv that are in this specified path, full.names --> tells R the right place to look 
tides_files

## now put it all together in the df and use previous tidyverse knowledge to summarize and create new columns 
maps_intertidaldata<-tides_files %>% ## tidal_files() is what is getting iteratred over everything 
  set_names()%>% # sets the id of each list to the file name
  map_df(read_csv,.id = "filename") %>% # map everything to a dataframe and put the id in a column called filename --> this is what is iterating 
  group_by(filename) %>%
  summarise(mean_temp = mean(Temp.C, na.rm = TRUE),
            mean_light = mean(Intensity.lux, na.rm = TRUE), 
            stdv_temp = sd(Temp.C, na.rm = TRUE), 
            stdv_light = sd(Intensity.lux, na.rm = TRUE)) %>% 
  mutate(filename = str_sub(filename, start= -7, end= -1))

maps_intertidaldata

```















